### 2024-06-28 paper_replication_buka.Rmd
paper replication with attempted extensions

---
title: "Replication of \"Anti-Immigrant Rhetoric and ICE Reporting Interest: Evidence
  from a Large-Scale Study of Web Search Data\" paper"
author: "Ubaydul Sami"
date: "2024-06-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(scales)
library(tidyverse)
library(knitr)
library(dplyr)
library(modelr)
library(scales)
library(readr)
```

```{r}

google_term_data <- read.csv('from_google_trends/google_3_terms_trend_data_04_20.csv', header = TRUE, sep= ",", skip = 2, col.names = c('ymd', 'immigr_report', 'immigr_crime', 'immigr_welfare'))

head(google_term_data)
```
# Figure 4 from paper
```{r}
google_term_data <- 
  google_term_data |>
  mutate(ymd = as.Date(paste0(ymd, "-01")))|>
  mutate(year = year(ymd), month = month(ymd)) |>
  mutate(president = ifelse(year<=2008, "bush", 
                      ifelse(year>2008 & year <= 2016, "obama", "trump")))
```

```{r}
google_term_data |> 
  ggplot(aes(x = ymd, y= immigr_report, colour = president)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Time Line") +
  ylab("Immigration Report Search Trend") +
  ggtitle("Immigration Report Search Trend on Google Trend Data")
```

```{r}
google_term_data |> 
  ggplot(aes(x = ymd, y= immigr_crime, colour = president)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Time Line") +
  ylab("Immigration Crime Search Trend") +
  ggtitle("Immigration Crime Search Trend on Google Trend Data")
```

```{r}
google_term_data |> 
  ggplot(aes(x = ymd, y= immigr_welfare, colour = president)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Time Line") +
  ylab("Immigration Welfare Search Trend") +
  ggtitle("Immigration Welfare Search Trend on Google Trend Data")
```

# Table 3 of the paper
```{r}
library(stargazer)
library(fastDummies)
```

```{r}
google_term_data <-
  google_term_data |>
  mutate(president = as.factor(president))

google_term_data_dummies <- dummy_cols(google_term_data, select_columns = "president", remove_first_dummy = FALSE)

head(google_term_data_dummies)
```

```{r}
model_crime <- lm(immigr_crime ~ (ymd + president_bush + president_trump + president_obama), data = google_term_data_dummies)
model_welfare <- lm(immigr_welfare ~ (ymd + president_bush + president_trump + president_obama), data = google_term_data_dummies)
model_report <- lm(immigr_report ~ (ymd + president_bush + president_trump + president_obama ), data = google_term_data_dummies)

```

```{r}
stargazer(model_crime, model_welfare, model_report,
          title = "Table 3 OLS",
          dep.var.labels = c("Crime", "Welfare", "Report"),

          out = "table3_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01")
```

# Figure 2 from paper
```{r}
library(stm)
load("TopicModel.RData")
document_topics <- make.dt(immigrFit, meta = out$meta)
topic_terms <- t(exp(immigrFit$beta$logbeta[[1]]))
rownames(topic_terms) <- out$vocab
colnames(topic_terms) <- sprintf("Topic%d", 1:ncol(topic_terms))
```

```{r}
document_topics_figure <- document_topics |>
  mutate(date = as.Date(date),
         year_month = ym(format(date, "%Y-%m")))
```

```{r}
# document_topics_figure |>
#   select(year_month, channel,duration, time) |>
#   group_by(year_month, channel, time) |>
#   summarise(air_count= sum(duration)) |> 
#   ggplot(aes(x = year_month, y = air_count, colour = channel)) +
#   geom_point()
```

```{r}

time_lines <- document_topics_figure |>
  select(time, year_month) |>
  group_by(time) |>
  summarise(year_month = min(year_month))

document_topics_figure |>
  select(year_month, channel,duration, time) |>
  group_by(year_month, channel, time) |>
  summarise(air_count= sum(duration)) |>
  ungroup() |> 
  group_by(channel, time) |> 
  ggplot(aes(x = year_month, y = air_count, color = channel, group = interaction(channel, time))) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year") +
  ylab("Num Monthly Immigration Segment") +
  ggtitle("Immigration news segments")

```

# Figure 3

```{r}
document_topics_figure <- document_topics_figure |>
  select(date, Topic1, Topic3, Topic13,year_month, channel, duration, time)

```

```{r}
document_topics_figure |>
  group_by(year_month, channel, time) |>
  summarise(topic1_seg = sum((Topic1 +Topic3)*duration )) |>
  ggplot(aes(x = year_month, y = topic1_seg, color = channel, group = interaction(channel, time))) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year") +
  ylab("Immigration Crime News Coverage") +
  ggtitle("Immigration news segments by Topic 'Immigration Crime'")
```

```{r}
document_topics_figure |>
  group_by(year_month, channel, time) |>
  summarise(topic1_seg = sum((Topic13)*duration )) |>
  ggplot(aes(x = year_month, y = topic1_seg, color = channel, group = interaction(channel, time))) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year") +
  ylab("Immigration Welfare News Coverage") +
  ggtitle("Immigration news segments by Topic 'Immigration Welfare'")
```
# Table 4 from paper:

```{r}

set.seed(1)

searches <-
  read_csv("C:/Users/buka/Documents/coursework/week4/immigrants.csv", skip = 3, col_names = c("month", "report", "crime", "welfare"))

adj <-
  read_csv("C:/Users/buka/Documents/immigrants/immigrant-news-2024-group-1/from_replication_files/gt_report_daily.csv") %>%
  rename(ymd = date)

month_model_data <-
  document_topics_figure %>%
  group_by(year_month) %>%
  summarize(
    segs = n(),
    crime_prop = sum(Topic1) + sum(Topic3),
    welfare_prop = sum(Topic13)
  ) %>%
  rename(ymd = year_month) %>%
  left_join(google_term_data_dummies)

day_model_data <-
  document_topics_figure %>%
  mutate(
    segs = n(),
    crime_prop = Topic1 + Topic3,
    welfare_prop = Topic13
  ) %>%
  rename(ymd = date) %>%
  left_join(adj) %>%
  mutate(president_trump = ifelse(ymd >= "2017-01-20", 1, 0))

model_month <- lm(immigr_report ~
              segs +
              crime_prop +
              welfare_prop +
              ymd +
              president_trump, 
            data = month_model_data)

model_day <- lm(search_adj ~
              segs +
              crime_prop +
              welfare_prop +
              ymd +
              president_trump, 
            data = day_model_data)

```

```{r}
stargazer(model_month,
          title = "Table 4 OLS Monthly",
          dep.var.labels = c("Report"),
          out = "table4_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01"
)

stargazer(model_day,
          title = "Table 4 OLS Daily",
          dep.var.labels = c("Report"),
          out = "table4_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01"
)
```

--- POSSIBLE EXTENTION QUESTIONS ---
do different news outlets have stronger or weaker effects on searches ?

maybe people want to move if they believe illegal immigrant population is high ? - (housing data ?)
  or how do reports about immigrant crime etc wffect people moving (topic 27, 3, 1)

do people really act with ice reportings ? is it reflected in anti-immigrant searches ? - (ice data)
  (test statement that "First, at least some searches for ‘how to report an immigrant’ will likely      result in an actual denunciation.")

which states have the highest suspicion of immigrants (highest searches) and which states did trump     win the majority ? is there a correlation ?

```{r}

trump_votes <-
  read_csv("elections_2016.csv", skip = 4, col_names = c("state", "", "", "votes", "", "", "state_total")) %>%
  select(state, votes, state_total) %>%
  head(-7) %>%
  rename(state_abb = state)

report_searches <-
  read_csv("report_popularity.csv", skip = 3, col_names = c("state", "searches")) %>%
  mutate(state_abb = state.abb[match(state, state.name)])
report_searches[1, 3] <- "DC"

data_by_state <-
  trump_votes %>%
  left_join(report_searches, by="state_abb") %>%
  mutate(votes = as.numeric(gsub(",", "", votes)), perc = votes / state_total, state_abb = as.factor(state_abb))

data_by_state %>%
  ggplot(aes(x = searches, y = perc)) +
  geom_point()

```

```{r}

searches_report <-
  read_csv("C:/Users/buka/Documents/immigrants/immigrant-news-2024-group-1/from_replication_files/gt_report_daily.csv")

preelections <-
  document_topics_figure %>%
  left_join(searches_report) %>%
  mutate(yr = year(date), total_news = Topic1 + Topic3 + Topic13) %>%
  filter(yr %in% c(2014, 2015, 2018, 2019)) %>%
  select(date, yr, Topic1, Topic3, Topic13, search_adj, total_news)

preelections %>%
  ggplot(aes(x = total_news, y= search_adj)) + 
  geom_point()

```