### 2024-06-28 paper_replication_sami.Rmd
after finish of draft

---
title: "Replication of \"Anti-Immigrant Rhetoric and ICE Reporting Interest: Evidence
  from a Large-Scale Study of Web Search Data\" paper"
author: "Ubaydul Sami and Buka Dikeocha"
date: "2024-06-24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(scales)
library(tidyverse)
library(knitr)
library(dplyr)
library(modelr)
library(scales)
library(readr)
```

```{r}

google_term_data <- read.csv('from_google_trends/google_3_terms_trend_data_04_20.csv', header = TRUE, sep= ",", skip = 2, col.names = c('ymd', 'immigr_report', 'immigr_crime', 'immigr_welfare'))

head(google_term_data)
```
# Figure 4 from paper
```{r}
google_term_data <- 
  google_term_data |>
  mutate(ymd = as.Date(paste0(ymd, "-01")))|>
  mutate(year = year(ymd), month = month(ymd)) |>
  mutate(president = ifelse(year<=2008, "bush", 
                      ifelse(year>2008 & year <= 2016, "obama", "trump")))
```

```{r}
google_term_data |> 
  ggplot(aes(x = ymd, y= immigr_report, colour = president)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Time Line") +
  ylab("Immigration Report Search Trend") +
  ggtitle("Immigration Report Search Trend on Google Trend Data")
```
```{r}
google_term_data |> 
  ggplot(aes(x = ymd, y= immigr_crime, colour = president)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Time Line") +
  ylab("Immigration Crime Search Trend") +
  ggtitle("Immigration Crime Search Trend on Google Trend Data")
```

```{r}
google_term_data |> 
  ggplot(aes(x = ymd, y= immigr_welfare, colour = president)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Time Line") +
  ylab("Immigration Welfare Search Trend") +
  ggtitle("Immigration Welfare Search Trend on Google Trend Data")
```

# Table 3 of the paper
```{r}
library(stargazer)
library(fastDummies)
```

```{r}
google_term_data <-
  google_term_data |>
  mutate(president = as.factor(president),
         president = relevel(president, ref = "obama"))

google_term_data_dummies <- dummy_cols(google_term_data, select_columns = "president", remove_first_dummy = FALSE)

head(google_term_data_dummies)
```
```{r}
model_crime <- lm(immigr_crime ~ (ymd + president_bush + president_trump + president_obama), data = google_term_data_dummies)
model_welfare <- lm(immigr_welfare ~ (ymd + president_bush + president_trump + president_obama), data = google_term_data_dummies)
model_report <- lm(immigr_report ~ (ymd + president_bush + president_trump + president_obama ), data = google_term_data_dummies)

```

```{r}
model_crime2 <- lm(immigr_crime ~ ymd + president, data = google_term_data_dummies)
model_welfare2 <- lm(immigr_welfare ~ (ymd + president), data = google_term_data_dummies)
model_report2 <- lm(immigr_report ~ (ymd + president), data = google_term_data_dummies)
```

# with dummies
```{r}
stargazer(model_crime, model_welfare, model_report,
          title = "Table 3 OLS",
          dep.var.labels = c("Crime", "Welfare", "Report"),

          out = "table3_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01" )
```
# with factored variable

```{r}
stargazer(model_crime2, model_welfare2, model_report2,
          title = "Table 3 OLS",
          dep.var.labels = c("Crime", "Welfare", "Report"),

          out = "table3_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01" )
```

# Figure 2 from paper
```{r}
library(stm)
load("from_replication_files/TopicModel.RData")
document_topics <- make.dt(immigrFit, meta = out$meta)
topic_terms <- t(exp(immigrFit$beta$logbeta[[1]]))
rownames(topic_terms) <- out$vocab
colnames(topic_terms) <- sprintf("Topic%d", 1:ncol(topic_terms))
```
```{r}
document_topics_figure <- document_topics |>
  mutate(date = as.Date(date),
         year_month = ym(format(date, "%Y-%m")))
```

```{r}
# document_topics_figure |>
#   select(year_month, channel,duration, time) |>
#   group_by(year_month, channel, time) |>
#   summarise(air_count= sum(duration)) |> 
#   ggplot(aes(x = year_month, y = air_count, colour = channel)) +
#   geom_point()
```

```{r}

time_lines <- document_topics_figure |>
  select(time, year_month) |>
  group_by(time) |>
  summarise(year_month = min(year_month))

document_topics_figure |>
  select(year_month, channel,duration, time) |>
  group_by(year_month, channel, time) |>
  summarise(air_count= sum(duration)) |>
  ungroup() |> 
  group_by(channel, time) |> 
  ggplot(aes(x = year_month, y = air_count, color = channel, group = interaction(channel, time))) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year") +
  ylab("Num Monthly Immigration Segment") +
  ggtitle("Immigration news segments")

```

# Figure 3

```{r}
document_topics_figure <- document_topics_figure |>
  select(date, Topic1, Topic3, Topic13,year_month, channel, duration, time)

```

```{r}
document_topics_figure |>
  group_by(year_month, channel, time) |>
  summarise(topic1_seg = sum((Topic1 +Topic3)*duration )) |>
  ggplot(aes(x = year_month, y = topic1_seg, color = channel, group = interaction(channel, time))) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year") +
  ylab("Immigration Crime News Coverage") +
  ggtitle("Immigration news segments by Topic 'Immigration Crime'")
```

```{r}
document_topics_figure |>
  group_by(year_month, channel, time) |>
  summarise(topic1_seg = sum((Topic13)*duration )) |>
  ggplot(aes(x = year_month, y = topic1_seg, color = channel, group = interaction(channel, time))) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year") +
  ylab("Immigration Welfare News Coverage") +
  ggtitle("Immigration news segments by Topic 'Immigration Welfare'")
```
# Table 4 from paper (by month):

```{r}
head(google_term_data_dummies)
```

```{r}
google_term_data_table4 <- google_term_data_dummies|> 
  mutate(month= as.factor(month), president_trump= as.factor(president_trump)) |>
  # filter(president_trump == 1) |>
  select(ymd, month, immigr_report, president_trump)

head(google_term_data_table4)
```
```{r}
head(document_topics_figure)
```
```{r}
document_topics_table4 <- 
  document_topics_figure |>
  mutate(crime = Topic1 + Topic3, welfare = Topic13, ymd = year_month) |>
  # select(crime, welfare, ymd) |>
  group_by(ymd) |>
  summarise(
    segs = n(), 
    crime_prop = mean(crime), 
    welfare_prop = mean(welfare))|> view()

head(document_topics_table4)
```
```{r}
df_table4_month <- left_join(document_topics_table4, google_term_data_table4, by = "ymd")

head(df_table4_month)
```

```{r}
set.seed (1)
model_report_table4_month <- lm(immigr_report ~ (segs  + crime_prop + welfare_prop + president_trump + ymd + month ), data = df_table4_month)
summary(model_report_table4_month)
```
```{r}
stargazer(model_report_table4_month,
          title = "Table 4 OLS",
          dep.var.labels = c("Report"),

          out = "table4_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01" )
```
# Table 4 from paper (by day):

```{r}
daily_report <- read.csv('from_replication_files/gt_report_daily.csv')
head(daily_report)

```

```{r}

```

```{r}
head(daily_report)
```

```{r}
daily_report_table4 <- 
  daily_report|> 
  mutate( date= as.Date(date), president_trump= ifelse(date< "2017-01-01", 0, 1), president_trump = as.factor(president_trump)) |>
  select(-X)

head(daily_report_table4)
```
```{r}
head(document_topics_figure)
```
```{r}
document_topics_table4_day <- 
  document_topics_figure |>
  mutate(crime = Topic1 + Topic3, welfare = Topic13) |>
  group_by(date) |>
  summarise(
    segs = n(), 
    crime_prop = mean(crime), 
    welfare_prop = mean(welfare))|> view()

head(document_topics_table4_day)
```
```{r}
df_table4_day <- left_join(document_topics_table4_day, daily_report_table4, by = "date")

df_table4_day <- df_table4_day |>
  mutate( month = month(date), day_of_week = weekdays(date), 
          day_of_week = as.factor(day_of_week), month = as.factor(month))

head(df_table4_day)
```

```{r}
set.seed (1)
model_report_table4_day <- lm(search ~ (segs  + crime_prop + welfare_prop + president_trump + date + day_of_week+ month ), data = df_table4_day)
summary(model_report_table4_day)
```
```{r}
stargazer(model_report_table4_day,
          title = "Table 4 OLS (per day)",
          dep.var.labels = c("Report"),

          out = "table4_ols.html", # Save to HTML file
          type = "text", # You can also use "html" or "latex"
          intercept.bottom = FALSE,
          omit.stat = c("ser", "adj.rsq", "f", "aic"),
          notes = "Note: ∗p<0.1; ∗∗p<0.05; ∗∗∗p<0.01" )
```
# Extension

## The paper shows increase in search of how to report of immigration violation but doesn't show if there were actual increase in number of reports / arrest/ deportation for immigration violation. For our research question we decided to find out if there was any increase of actual number of reports / arrest/ deportation for immigration violation.

**data source :** https://www.dhs.gov/ohss/topics/immigration/enforcement-and-legal-processes-monthly-tables

**link to data file:** https://onedrive.live.com/edit?id=6779012659AEA485%21120&resid=6779012659AEA485%21120&ithint=file%2Cxlsx&wdPreviousSession=2e2e85d3-ac39-43da-899d-f01ad6e066a7&wdo=2&cid=6779012659aea485

```{r}

```

```{r}
google_term_data_report_extend <- 
  google_term_data_dummies |>
  filter(ymd >= '2014-01-01') |> 
  rename( year_month = ymd) |>
  select(year_month, year, month, president, immigr_report ) |>
  left_join(time_lines, by = "year_month") |>
  fill(time, .direction = c("down")) |> view()

head(google_term_data_report_extend)
```

```{r}
google_term_data_report_extend |>
  ggplot(aes(x = year_month, y = immigr_report, colour = time)) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +
  xlab("Year-month") +
  ylab("Immigration Reports search count") +
  ggtitle("Immigration report search")
```

```{r}

ice <-
  read_csv("from_google_trends/ice_reportings.csv", skip = 4, col_names = c("year", "month", "reports", "reports_convicted", "", "", "", "reports_non_criminal", "", "")) %>%
  select(c(1, 2, 3, 4, 8)) %>%
  fill(year, .direction = "down") %>% 
  filter(month != "Total") %>%
  mutate(
    month = match(month, month.name),
    date = make_date(year, month)
  )

```

```{r}
ice_report_w_election <- 
  ice |>
  filter(date <= '2020-01-01') |> 
  rename( year_month = date) |>
  # select(year_month, year, month, president, immigr_report ) |>
  arrange(year_month) |>
  left_join(time_lines, by = "year_month") |>
  fill(time, .direction = c("down")) |>  view()

head(ice_report_w_election)
```

```{r}
ice_report_w_election %>%
  ggplot(aes(x = year_month, y = reports, colour = time)) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +

  xlab("Year-month") +

  ylab("Total Immigration Arrest") +

  ggtitle("Total Immigration Arrest by Year")

ice_report_w_election %>%
  ggplot(aes(x = year_month, y = reports_convicted, colour = time)) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +

  xlab("Year-month") +

  ylab("Convicted Immigration Arrest") +

  ggtitle("Convicted Immigration Arrest by Year")

ice_report_w_election %>%
  ggplot(aes(x = year_month, y = reports_non_criminal, colour = time)) + 
  geom_point() + geom_smooth( se=FALSE) +
  geom_vline(data = time_lines, aes(xintercept = as.numeric(year_month)), linetype = "dotted", color = "black") +

  xlab("Year-month") +

  ylab("Non-criminal Immigration Arrest") +

  ggtitle("Non-Criminal Immigration Arrest by Year")

```